`localhierarchy`: an R package to facilitate fitting of global and local Bayesian hierarchical models

### 2025-09-30 paper/paper.qmd
misc paper editing, refs

---
title: "`localhierarchy`: an R package to facilitate fitting of global and local Bayesian hierarchical models"
author: "Leontine, Shauna, Evan, Herb"
format: 
  pdf:
    latex-engine: pdflatex    # lualatex or xelatex or pdflatex
    include-in-header: 
      - file: preamble.tex 
editor: visual
execute:
  cache: true

@@

  This article presents a practical introduction to the package for the applied user and illustrates the package’s functionality through examples for national and subnational estimation.
date: last-modified
bibliography: references.bib
number-sections: true
---

@@

In global health estimation, Bayesian hierarchical models have become a cornerstone methodology, enabling the integration of diverse data sources to produce reliable estimates across multiple populations and regions. For example, they are used to estimate family planning indicators for the FP2030 initiative (ref), low birth weight (ref), and maternal mortality by the United Nations Interagency Group (ref). In these efforts, national or subnational estimates are often required for population-periods with limited data availability. Bayesian hierarchical models—whether used directly or for specific components within larger modeling frameworks—are valuable in these contexts because they facilitate the sharing of information across populations. As a result, global modeling exercises typically fit Bayesian hierarchical models to comprehensive global databases, producing national-level estimates for all countries worldwide.

@@

To facilitate local-level analysis, it is often desirable to use models that are informed by global models but can be fitted to just a subset of the data, such as data from a single country. We refer to these as local models: models derived from global hierarchical models but adapted to be fitted using data from only one population. Beyond allowing model fitting to smaller, specific datasets, local models offer practical advantages such as faster computation, making it feasible for users to perform analyses quickly on their own computers. Additionally, local models enable estimates for different populations to be updated independently, for example, when new data become available. This approach has been used successfully in global estimation exercises for family planning and maternal mortality (ref, ref).

@@

```{r}
#| fig-height: 9
#| fig-width: 7
#| echo: false
#| label: fig-localfits
#| fig-cap: "Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in Benin (BEN), Gambia (GMB), and Ghana (GHA). Graphs show mCPR, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals."
plots <- readRDS("outputfpet_localfits.rds")
plots[[1]] /  plots[[2]]/ plots[[3]] 
```

At its core, the Bayesian FPET model employs transition models for demand and demand satisfied to characterize the long-term trajectories of family planning indicators over time within a population, illustrated in @fig-transition. The transition model describes changes in an indicator in terms of three principal parameters: the rate of change, the timing of the transition, and the asymptotic level (the maximum attainable level for an indicator). To allow for population-specific variation in how family planning indicators evolve, the parameters of the transition model are population-specific.

![Illustration of transition model used in the modeling of family planning indicators. TO DO (if kept): make specific to this text, ie name asymptote, level](fig/transition.png){#fig-transition}

@@

```{r, echo = FALSE}
plots <- readRDS("outputfpet_params.rds")
```

```{r}
#| fig-height: 9
#| fig-width: 7
#| echo: false
#| label: fig-level
#| fig-cap: "Estimates of the level of demand for countries in Western Africa and for higher levels used in the hierarchical model (smaller world regions and major world regions). Point estimates and 95% credible intervals are shown. TO DO: fix formatting, show country names"
plots$plot_iso_omega /  plots$plot_reg_omega[[3]]/ plots$plot_reg_omega[[2]]
```

```{r}
#| fig-height: 9
#| fig-width: 7
#| echo: false
#| label: fig-asympt
#| fig-cap: "Estimates of the asymptote for demand for countries in Western Africa and for higher levels used in the hierarchical model (major world regions). Point estimates and 95% credible intervals are shown. TO DO: fix formatting, show country names"
plots$plot_iso_ptilde /  plots$plot_reg_ptilde[[2]] 
```

The workflow used in FPET to produce country-level estimates is as follows. Firstly, the full Bayesian model is fitted to a global data base of survey data. Data base compilation is carried out by the United Nations Population Division (ref) and the Track20 project (ref). The global model fit provides information on model parameters, including hierarchical parameters, e.g., estimates of asymptotes. Subsequently, a local model can be used to produce FP estimates based on the data from one country, using information from the global model fit. The technical details for this approach are given in the next section. We return to the application of estimating family planning indicators in Section 4.2.

@@

Suppose we want to estimate a population parameter $\mu_r^{(l)}$ of interest, referring to a population health indicator for region $r$ at level $l$ of the hierarchy. For example, the indicator could refer to some population health indicator such as service coverage or morbidity or mortality-related outcomes. For a given level $l$, index $r$ refers to the population in a particular region indexed by $r$, for example, a subnational region, a country, or a group of countries. We assume that $\mu$ is unconstrained (possibly after a transformation has been applied) and consider the following general hierarchical model setup, see Figure 6: 
\begin{align*}
  \mu_{r}^{(l)} \mid \mu^{(l-1)}_{r^\prime}, \sigma_l} &\sim N(\mu^{(l-1)}_{r^\prime}, \sigma_l^2), \text{ for } l > 0,
\end{align*} where $\mu^{(l-1)}_{r^\prime}$ is the parameter value at one level higher in the hierarchy for the region $r^\prime$ that contains unit $r$. The parameter $\sigma_l$ for $l > 0$ describes variability of values of the parameter $\mu$ across geographic units at level $l$ of the hierarchy. At the top level of the hierarchy, $l = 0$, only a single parameter is estimated.

@@

\begin{alignat}{2}
  \mu_c^{(\country)}^{(\country)} &\mid \mu_{s[c]}^{(\subregion)}, \sigma_{ \country}^2 &&\sim N(\mu_{s[c]}^{(\subregion)}, \sigma_{ \country}^2), \label{eq-countryhier} \\
  \mu_s^{(\subregion)} &\mid \mu_{w[s]}^{(\region)}, \sigma_{ \subregion}^2 &&\sim N(\mu_{w[s]}^{(\region)}, \sigma_{ \subregion}^2), \nonumber \\
  \mu_w^{(\region)} &\mid \mu^{(\glob)}, \sigma_{ \region}^2 &&\sim N(\mu^{(\glob)}, \sigma_{ \region}^2), \nonumber
\end{alignat} where $c =1,\dots,C$ refers to the country index, $s[c] = 1, .., S$ to the subregion index of country $c$, and $w[s[c]] = 1, .., W$ to the region index of subregion $s[c]$. The parameter $\mu^{(\glob)}$ is the global mean for the parameter of interest, e.g., a global mean for service coverage or mortality. The parameters $\sigma_{ \country}$, $\sigma_{ \subregion}$, and $\sigma_{ \region}$ describe variability in the parameter of interest across countries, subregions, and regions, respectively.

@@

\[to finish, LA took mostly from fpemlocal paper\] `localhierarchy` is a publicly available R package stored on Github. For usage, an R (R Core Team, 2020) installation (≥) and a cmdStan (xxx) installation (≥xxx) are required. R can be downloaded on CRAN. Stan install \[add info\] xxx . The `localhierarchy` package can be installed with one of two methods: (1) direct install from Github repository using the `install_ github()` API from the devtools package (Wickham et al., 2020); (2) install from the package binary using the base R function `install.package()`. Package dependencies are listed in the package DESCRIPTION file and will be automatically installed upon installing the main package. There are no minimum RAM, CPU, or HARDDRIVE requirements apart from what is necessary to store model runs, which varies case-by-case.

@@

\node[level1] (mu0) {$\mu^{(1)}_1$};

@@

% keep same colors so now color of level 1 at top 
\node[level2] (mu0) {$\mu^{(2)}_1$};

@@

For family planning estimation, a local model fitting approach was motivated by the need to enable country monitoring and evaluation (M&E) officers to produce estimates for their countries, as part of the Track20 project (ref). The Track20 project works to build the capacity of countries to generate and use data to inform their programming, improve impact, and accelerate progress toward their FP goals. One of the pillars of this approach is the cultivation of a network of M&E officers dedicated to increasing the quality and use of FP data. We developed the family planning estimation tool (FPET) as a local version of a global family planning model, for use by the M&E officers.

For local estimation of family planning indicators, the Bayesian model described in section 2 is used. Specifically, hierarchical models are used for estimation of transition model parameters. The workflow used for estimating family planning indicators at the national level is based on model fitting in several steps, including the estimation of long-term trends in the first step, the estimation of all model parameters related to modern use in a second step, the addition of information on the use of traditional methods of contraception in a third step, followed by local model fitting to produce estimates for all indicators of interest. The `localhierarchy` functionality is incorporated in the family planning modeling code base such that all steps are run with the same code base, and parameters can be passed from one model fit to another. Full details on this workflow and model specification are described elsewhere (ref). As an illustration, @fig-allfits shows global and local fits for modern use, demand, and demand satisfied for selected countries in Western Africa. The global fit in red corresponds closely to the local fits, shown in green.

```{r}
#| fig-height: 9
#| fig-width: 7
#| echo: false
#| label: fig-allfits
#| fig-cap: "Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in in Benin (BEN), Gambia (GMB), and Ghana (GHA). FPET estimates are shown in in colored lines (red for global, green for local), with shaded areas representing 90% credible intervals. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance."
plots <- readRDS("outputfpet_localglobalfits.rds")
plots[[1]] /  plots[[2]]/ plots[[3]] # Stack three plots vertically
```

@@

The figures below show illustrations of regional fits for Benin. The subnational model is fitted to all regional data from the country. @fig-subnatoverview illustrates fits in selected regions and differences across regions in 2024.

```{r}
#| fig-height: 9
#| fig-width: 7
#| echo: false
#| label: fig-subnatfits
#| fig-cap: "Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in subnational regions in Benin. Graphs show mCPR, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals. "
plots <- readRDS("outputfpet_plotssubnat_BEN.rds")
plots[["Oueme"]]$married / plots[["Collines"]]$married/ plots[["Donga"]]$married
```

```{r}
#| echo: false
#| label: fig-subnatoverview
#| fig-cap: "Regional estimates for Benin for modern contraceptive use in 2024. Point estimates and 95% credible intervals shown."
plots[["comp"]] 
```

@@

We have used the `localHierarchy` functionality in different projects. We discussed the estimation of family planning indicators, where country M&E officers use the local model to produce estimates at the national and subnational level. In this area, M&E officers have successfully used local models to produce national estimates since 2014. Availability of a local modeling tool has enabled users to do additional analyses and check sensitivity of estimates to specific data sources or data quality assumptions. In addition to fitting the model to survey data, users of the local family planning model can also incorporate additional data from routine information systems (e.g., the number of contraceptive commodities supplied, add REF). These use cases illustrate the value of local models to enable local analysis.

More recently, we have used the `localHierarchy` approach and functionality for estimating coverage indicators for the Countdown to 2030 initiative (ref). By developing the `localHierarchy` package and providing use cases in this paper, we hope that this work can be useful for other global health modelers, to consider local modeling in their own work.

@@

>>>>

---
title: ""
author: ""
format:
  pdf:
    latex-engine: pdflatex    # lualatex or xelatex or pdflatex
    include-before-body: title.tex
    include-in-header: 
      - file: preamble.tex 
editor: visual
execute:
  cache: true

@@

  This article presents a practical introduction to the package for the applied user and illustrates the package’s functionality through examples for national and subnational estimation.
bibliography: bibliography.bib
number-sections: true
---

@@

In global health estimation, Bayesian hierarchical models have become a cornerstone methodology, enabling the integration of diverse data sources to produce reliable estimates across multiple populations and regions. For example, they are used to estimate family planning indicators for the FP2030 initiative [@alkema_2024evol], incidence of low birth weight [@okwaraji_national_2024], abortion and unintended pregnancies rates [@bearak_unintended_2020], and maternal mortality by the United Nations Interagency Group [@peterson_bayesian_2024]. In these efforts, national or subnational estimates are often required for population-periods with limited data availability. Bayesian hierarchical models—whether used directly or for specific components within larger modeling frameworks—are valuable in these contexts because they facilitate the sharing of information across populations. As a result, global modeling exercises typically fit Bayesian hierarchical models to comprehensive global databases, producing national-level estimates for all countries worldwide.

@@

To facilitate local-level analysis, it is often desirable to use models that are informed by global models but can be fitted to just a subset of the data, such as data from a single country. We refer to these as local models: models derived from global hierarchical models but adapted to be fitted using data from only one population. Beyond allowing model fitting to smaller, specific datasets, local models offer practical advantages such as faster computation, making it feasible for users to perform analyses quickly on their own computers. Additionally, local models enable estimates for different populations to be updated independently, for example, when new data become available. This approach has been used successfully in global estimation exercises for family planning and maternal mortality [@alkema_2024evol; @peterson_bayesian_2024].

@@

![Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in Benin (BEN), Gambia (GMB), and Ghana (GHA). Graphs show mCPR, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals.](fig/fp_local.jpg){#fig-localfits}

<!-- ```{r} -->

<!-- #| fig-height: 9 -->

<!-- #| fig-width: 7 -->

<!-- #| echo: false -->

<!-- #| label: fig-localfits -->

<!-- #| fig-cap: "Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in Benin (BEN), Gambia (GMB), and Ghana (GHA). Graphs show mCPR, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals." -->

<!-- plots <- readRDS("outputfpet_localfits.rds") -->

<!-- plots[[1]] /  plots[[2]]/ plots[[3]]  -->

<!-- ``` -->

At its core, the Bayesian FPET model employs transition models for demand and demand satisfied to characterize the long-term trajectories of family planning indicators over time within a population [@susmann_flexible_2025], illustrated in @fig-transition. The transition model describes changes in an indicator in terms of three principal parameters: the rate of change, the timing of the transition, and the asymptotic level (the maximum attainable level for an indicator). To allow for population-specific variation in how family planning indicators evolve, the parameters of the transition model are population-specific.

![Schematic illustration of transition model used in the modeling of family planning indicators, modified from @susmann_flexible_2025. The transition model describes changes in an indicator (here $\eta_{c,t}$ in terms of three principal parameters: the rate of change, the level in a reference year (here labeled with $\Omega_c$), and the asymptotic level (the maximum attainable level for an indicator (here labeled with $\lambda_c$).](fig/transition.png){#fig-transition width="476"}

@@

<!-- ```{r, echo = FALSE} -->

<!-- plots <- readRDS("outputfpet_params.rds") -->

<!-- ``` -->

<!-- ```{r} -->

<!-- #| fig-height: 9 -->

<!-- #| fig-width: 7 -->

<!-- #| echo: false -->

<!-- #| label: fig-level -->

<!-- #| fig-cap: "Estimates of the level of demand for countries in Western Africa and for higher levels used in the hierarchical model (smaller world regions and major world regions). Point estimates and 95% credible intervals are shown. " -->

<!-- plots$plot_reg_omega[[2]]/ plots$plot_reg_omega[[3]]/  plots$plot_iso_omega -->

<!-- ``` -->

<!-- ```{r} -->

<!-- #| fig-height: 9 -->

<!-- #| fig-width: 7 -->

<!-- #| echo: false -->

<!-- #| label: fig-asympt -->

<!-- #| fig-cap: "Estimates of the asymptote for demand for countries in Western Africa and for higher levels used in the hierarchical model (major world regions). Point estimates and 95% credible intervals are shown. " -->

<!-- plots$plot_reg_ptilde[[2]]  / plots$plot_iso_ptilde -->

<!-- ``` -->

![Estimates of the level of demand for countries in Western Africa and for higher levels used in the hierarchical model (smaller world regions and major world regions). Point estimates and 95% credible intervals are shown.](fig/hier_level.jpg){#fig-level width="685"}

![Estimates of the asymptote for demand for countries in Western Africa and for higher levels used in the hierarchical model (major world regions). Point estimates and 95% credible intervals are shown.](fig/hier_asymptote.jpg){#fig-asympt}

The workflow used in FPET to produce country-level estimates is as follows. Firstly, the full Bayesian model is fitted to a global data base of survey data. Data base compilation is carried out by the United Nations Population Division (see [UNPD world contraceptive use website](https://www.un.org/development/desa/pd/world-contraceptive-use)) and the Track20 project (see [www.track20.org](https://www.track20.org/)). The global model fit provides information on model parameters, including hierarchical parameters, e.g., estimates of asymptotes. Subsequently, a local model can be used to produce FP estimates based on the data from one country, using information from the global model fit. The technical details for this approach are given in the next section. We return to the application of estimating family planning indicators in Section 4.2.

@@

Suppose we want to estimate a population parameter $\mu_r^{(l)}$ of interest, referring to a population health indicator for region $r$ at level $l$ of the hierarchy. For example, the indicator could refer to some population health indicator such as service coverage or morbidity or mortality-related outcomes. For a given level $l$, index $r$ refers to the population in a particular region indexed by $r$, for example, a subnational region, a country, or a group of countries. We assume that $\mu$ is unconstrained (possibly after a transformation has been applied) and consider the following general hierarchical model setup, see Figure 6: \begin{align*}
  \mu_{r}^{(l)} \mid \mu^{(l-1)}_{r^\prime}, \sigma_l &\sim N(\mu^{(l-1)}_{r^\prime}, \sigma_l^2), \text{ for } l > 0,
\end{align*} where $\mu^{(l-1)}_{r^\prime}$ is the parameter value at one level higher in the hierarchy for the region $r^\prime$ that contains unit $r$. The parameter $\sigma_l$ for $l > 0$ describes variability of values of the parameter $\mu$ across geographic units at level $l$ of the hierarchy. At the top level of the hierarchy, $l = 0$, only a single parameter is estimated.

@@

\begin{alignat}{2}
  \mu_c^{(\country)} &\mid \mu_{s[c]}^{(\subregion)}, \sigma_{ \country}^2 &&\sim N(\mu_{s[c]}^{(\subregion)}, \sigma_{ \country}^2), \label{eq-countryhier} \\
  \mu_s^{(\subregion)} &\mid \mu_{w[s]}^{(\region)}, \sigma_{ \subregion}^2 &&\sim N(\mu_{w[s]}^{(\region)}, \sigma_{\subregion}^2), \nonumber \\
  \mu_w^{(\region)} &\mid \mu^{(\glob)}, \sigma_{ \region}^2 &&\sim N(\mu^{(\glob)}, \sigma_{ \region}^2), \nonumber
\end{alignat} where $c =1,\dots,C$ refers to the country index, $s[c] = 1, .., S$ to the subregion index of country $c$, and $w[s[c]] = 1, .., W$ to the region index of subregion $s[c]$. The parameter $\mu^{(\glob)}$ is the global mean for the parameter of interest, e.g., a global mean for service coverage or mortality. The parameters $\sigma_{ \country}$, $\sigma_{ \subregion}$, and $\sigma_{ \region}$ describe variability in the parameter of interest across countries, subregions, and regions, respectively.

@@

`localhierarchy` is a publicly available R package stored on Github. For usage, installation of R (≥4.5.1), Stan (≥2.37.0), and CmdStanR (≥ 0.9.0) are required [@WeloveR; @stan2025; @cmdstanr_2025]. The `localhierarchy` package can be installed from the Github repository. Package dependencies are listed in the package DESCRIPTION file and will be automatically installed upon installing the main package. There are no minimum RAM, CPU, or HARDDRIVE requirements apart from what is necessary to store model runs, which varies case-by-case.

<!-- The `localhierarchy` package can be installed with one of two methods: (1) direct install from Github repository using the `install_ github()` API from the devtools package (Wickham et al., 2020); (2) install from the package binary using the base R function `install.package()`. Package dependencies are listed in the package DESCRIPTION file and will be automatically installed upon installing the main package. There are no minimum RAM, CPU, or HARDDRIVE requirements apart from what is necessary to store model runs, which varies case-by-case. -->

@@

\node[level1] (mu0) {$\mu^{(\country)}_1$};

@@

% keep same colors so now color of level 1 at top 
\node[level2] (mu0) {$\mu^{(\subnat)}_1$};

@@

For family planning estimation, a local model fitting approach was motivated by the need to enable country monitoring and evaluation (M&E) officers to produce estimates for their countries, as part of the Track20 project [@alkema_2024evol]. The Track20 project works to build the capacity of countries to generate and use data to inform their programming, improve impact, and accelerate progress toward their FP goals. One of the pillars of this approach is the cultivation of a network of M&E officers dedicated to increasing the quality and use of FP data. We developed the family planning estimation tool (FPET) as a local version of a global family planning model, for use by the M&E officers.

For local estimation of family planning indicators, the Bayesian model described in section 2 is used. Specifically, hierarchical models are used for estimation of transition model parameters. The workflow used for estimating family planning indicators at the national level is based on model fitting in several steps, including the estimation of long-term trends in the first step, the estimation of all model parameters related to modern use in a second step, the addition of information on the use of traditional methods of contraception in a third step, followed by local model fitting to produce estimates for all indicators of interest. The `localhierarchy` functionality is incorporated in the family planning modeling code base such that all steps are run with the same code base, and parameters can be passed from one model fit to another. Full details on this workflow and model specification are described elsewhere [@alkema_2024evol]. As an illustration, @fig-allfits shows global and local fits for modern use, demand, and demand satisfied for selected countries in Western Africa. The global fit in red corresponds closely to the local fits, shown in green.

![Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in in Benin (BEN), Gambia (GMB), and Ghana (GHA). FPET estimates are shown in in colored lines (red for global, green for local), with shaded areas representing 90% credible intervals. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance.](fig/fpet_localglobal.jpg){#fig-allfits}

<!-- ```{r} -->

<!-- #| fig-height: 9 -->

<!-- #| fig-width: 7 -->

<!-- #| echo: false -->

<!-- #| label: fig-allfits -->

<!-- #| fig-cap: "Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in in Benin (BEN), Gambia (GMB), and Ghana (GHA). FPET estimates are shown in in colored lines (red for global, green for local), with shaded areas representing 90% credible intervals. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance." -->

<!-- plots <- readRDS("outputfpet_localglobalfits.rds") -->

<!-- plots[[1]] /  plots[[2]]/ plots[[3]] # Stack three plots vertically -->

<!-- ``` -->

@@

Figure @fig-subnatfits show illustrations of regional fits for Benin. The subnational model is fitted to all regional data from the country. @fig-subnatoverview illustrates fits in selected regions and differences across regions in 2024.

![Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in subnational regions in Benin. Graphs show mCPR, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals.](fig/fpet_subnat.jpg){#fig-subnatfits}

![Regional estimates for Benin for modern contraceptive use in 2024. Point estimates and 95% credible intervals shown.](fig/subnat_overview.jpg){#fig-subnatoverview}

<!-- ```{r} -->

<!-- #| fig-height: 9 -->

<!-- #| fig-width: 7 -->

<!-- #| echo: false -->

<!-- #| label: fig-subnatfits -->

<!-- #| fig-cap: "Data, estimates, and forecasts from 1970 until 2030 for FP indicators for married women in subnational regions in Benin. Graphs show mCPR, unmet need for modern methods, demand, and demand satisfied with modern methods. Survey data are shown in colored dots with 95% confidence intervals that reflect total error variance. FPET estimates are shown in black, with grey shaded areas representing 90% credible intervals. " -->

<!-- plots <- readRDS("outputfpet_plotssubnat_BEN.rds") -->

<!-- plots[["Oueme"]]$married / plots[["Collines"]]$married/ plots[["Donga"]]$married -->

<!-- ``` -->

<!-- ```{r} -->

<!-- #| echo: false -->

<!-- #| label: fig-subnatoverview -->

<!-- #| fig-cap: "Regional estimates for Benin for modern contraceptive use in 2024. Point estimates and 95% credible intervals shown." -->

<!-- plots[["comp"]]  -->

<!-- ``` -->

@@

We have used the `localHierarchy` functionality in different projects. We discussed the estimation of family planning indicators, where country M&E officers use the local model to produce estimates at the national and subnational level. In this area, M&E officers have successfully used local models to produce national estimates since 2014. Availability of a local modeling tool has enabled users to do additional analyses and check sensitivity of estimates to specific data sources or data quality assumptions. In addition to fitting the model to survey data, users of the local family planning model can also incorporate additional data from routine information systems (e.g., the number of contraceptive commodities supplied, see [@mooney_emu]). These use cases illustrate the value of local models to enable local analysis.

More recently, we have used the `localHierarchy` approach and functionality for estimating coverage indicators for the *Countdown to 2030 for Women’s, Children’s and Adolescents’ Health initiative.* This initiative focuses on “tracking progress of life-saving interventions for Reproductive, Maternal, Newborn, Child and Adolescent Health and Nutrition” in low- and middle-income countries, see [www. countdown2030.org](https://www.%20countdown2030.org/). By developing the `localHierarchy` package and providing use cases in this paper, we hope that this work can be useful for other global health modelers, to consider local modeling in their own work.

@@

# References
